Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-03-12 draft.tex
Fixed references (journals, publishers, names, ...)

\item The approximate multivariate Gaussian distribution of $\vecest{S}$
    allows easier and more precise transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \cite{Altmannshofer:2013foa,
Descotes-Genon:2013wba,Beaujean:2013soa},
  or for more precise predictions of optimized observables; see e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $N \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

@@

The PDF for the decay --- in the presence of Standard-Model operators and their chirality-flipped counter parts --- reads \cite{Boeer:2014xx}
\begin{equation}
\begin{aligned}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3)
        & = \frac{(\rmdx{\Gamma} /\, \rmdx{q^2})^{-1} \rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}}\\
        & = \sum_i S_i f_i(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{aligned}
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and
$N\pi$ systems, respectively, and $\theta_3 = \phi$ denotes the azimuthal angle.
The index $i$ should be interpreted as a multi-index, $i \equiv (l_1, l_2, m)$,
where $0 \leq l_1 \leq 2$ and $0 \leq l_2 \leq 1$ denote the total angular momentum in the
dilepton and the $N\pi$ system, respectively, and $-1 \leq m \leq 1$
is the third component of either of the angular momenta.\\

@@

\begin{widetext}
The correspondence between out choice of angular observables in the angular momentum basis, and the angular
observables as defined in reference \cite{Boeer:2014xx} reads
\begin{equation}
\begin{aligned}
    8\pi S_{0, 0,  0} & = 1\,,                                                             &
    8\pi S_{0, 1,  0} & = \frac{K_{2cc} + 2 K_{2ss}}{\rmdx\Gamma/\rmdx{q^2}}\,,
\end{aligned}
\end{equation}
and
\begin{equation}
\begin{aligned}
%
                      &                                                                    &
    8\pi S_{1, 1, -1} & = \frac{6 K_{4s}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       \\
%
    8\pi S_{1, 0,  0} & = \frac{3 K_{1c}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    8\pi S_{1, 1,  0} & = \frac{3 K_{2c}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       \\
%
                      &                                                                    &
    8\pi S_{1, 1, +1} & = \frac{6 K_{3s}}{\rmdx\Gamma/\rmdx{q^2}}\,,
\end{aligned}
\end{equation}
and
\begin{equation}
\begin{aligned}
%
                      &                                                                    &
    8\pi S_{2, 1, -1} & = \frac{2 \sqrt{3} K_{4sc}}{\rmdx\Gamma/\rmdx{q^2}}\,,             \\
%
    8\pi S_{2, 0,  0} & = \frac{2(K_{1cc} - K_{1ss})}{\rmdx\Gamma/\rmdx{q^2}}\,,           &
    8\pi S_{2, 1,  0} & = \frac{2(K_{2cc} - K_{2ss})}{\rmdx\Gamma/\rmdx{q^2}}\,,           \\
%
                      &                                                                    &
    8\pi S_{2, 1, +1} & = \frac{2 \sqrt{3} K_{3sc}}{\rmdx\Gamma/\rmdx{q^2}}\,,
\end{aligned}
\end{equation}
where the decay width is
\begin{equation}
    \frac{\rmdx{\Gamma}} {\rmdx{q^2}} = 2 K_{1ss} + K_{1cc}\,.
\end{equation}

>>>>

\item The approximate multivariate Gaussian distribution of $\vecest{S}$
    allows easier and more precise transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \cite{Altmannshofer:2013foa,
Descotes-Genon:2013wba,Beaujean:2013soa},
  or for more precise predictions of optimized observables; see e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boer:2014kda} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $N \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

@@

The PDF for the decay --- in the presence of Standard-Model operators and their chirality-flipped counter parts --- reads \cite{Boer:2014kda}
\begin{equation}
\begin{aligned}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3)
        & = \frac{(\rmdx{\Gamma} /\, \rmdx{q^2})^{-1} \rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}}\\
        & = \sum_i S_i f_i(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{aligned}
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and
$N\pi$ systems, respectively, and $\theta_3 = \phi$ denotes the azimuthal angle.
The index $i$ should be interpreted as a multi-index, $i \equiv (l_1, l_2, m)$,
where $0 \leq l_1 \leq 2$ and $0 \leq l_2 \leq 1$ denote the total angular momentum in the
dilepton and the $N\pi$ system, respectively, and $-1 \leq m \leq 1$
is the third component of either of the angular momenta.\\

@@

\begin{widetext}
The correspondence between out choice of angular observables in the angular momentum basis, and the angular
observables as defined in reference \cite{Boer:2014kda} reads
\begin{equation}
\begin{aligned}
    8\pi S_{0, 0,  0} & = 1\,,                                                             &
    8\pi S_{0, 1,  0} & = \frac{K_{2cc} + 2 K_{2ss}}{\rmdx\Gamma/\rmdx{q^2}}\,,
\end{aligned}
\end{equation}
and
\begin{equation}
\begin{aligned}
%
                      &                                                                    &
    8\pi S_{1, 1, -1} & = \frac{6 K_{4s}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       \\
%
    8\pi S_{1, 0,  0} & = \frac{3 K_{1c}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    8\pi S_{1, 1,  0} & = \frac{3 K_{2c}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       \\
%
                      &                                                                    &
    8\pi S_{1, 1, +1} & = \frac{6 K_{3s}}{\rmdx\Gamma/\rmdx{q^2}}\,,
\end{aligned}
\end{equation}
and
\begin{equation}
\begin{aligned}
%
                      &                                                                    &
    8\pi S_{2, 1, -1} & = \frac{2 \sqrt{3} K_{4sc}}{\rmdx\Gamma/\rmdx{q^2}}\,,             \\
%
    8\pi S_{2, 0,  0} & = \frac{2(K_{1cc} - K_{1ss})}{\rmdx\Gamma/\rmdx{q^2}}\,,           &
    8\pi S_{2, 1,  0} & = \frac{2(K_{2cc} - K_{2ss})}{\rmdx\Gamma/\rmdx{q^2}}\,,           \\
%
                      &                                                                    &
    8\pi S_{2, 1, +1} & = \frac{2 \sqrt{3} K_{3sc}}{\rmdx\Gamma/\rmdx{q^2}}\,,
\end{aligned}
\end{equation}
where the decay width is
\begin{equation}
    \frac{\rmdx{\Gamma}} {\rmdx{q^2}} = 2 K_{1ss} + K_{1cc}\,.
\end{equation}
